---
title: "CAB Meeting Times"
author: "Stevie Pederson"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE,
  fig.height = 9, fig.width = 10
)
```

```{r packages}
library(timechange)
library(tidyverse)
library(ggtext)
library(pander)
panderOptions("missing", "")
panderOptions("table.split.table", Inf)
```

```{r draw-schedule}
draw_schedule <- function(
    dates, people, bus_hours = c(8, 18), upr = 23, lwr = 5,
    exec = c("Stevie", "Lori", "Janani"), ...
) {
  exec_regexp <- paste0("(", paste(exec, collapse = "|"), ")")
  people %>%
    lapply(
      \(x) mutate(
        dates,
        tz = x,
        local = time_at_tz(utc, tz = x),
        time =  format(local, "%H:%M"),
        time_dbl =  hour(local) + minute(local)/60
      )
    ) %>%
    bind_rows(.id = "member") %>%
    mutate(
      member = fct(
        member,
        levels = people %>%
          lapply(\(x) now(tzone = x)) %>%
          lapply(format, "%z") %>%
          map_dbl(as.numeric) %>%
          sort() %>%
          names()
      ) %>% 
        fct_relabel(\(x) str_replace_all(x, exec_regexp, "**\\1**"))
    ) %>%
    arrange(member) %>%
    mutate(
      tz = tz %>% str_replace_all("(/)(.+)", "\n(\\2)") %>% fct_inorder(),
      status = case_when(
        time_dbl >= upr | time_dbl <= lwr ~ "Horrible",
        time_dbl >= max(bus_hours) | time_dbl < min(bus_hours) ~ "Outside Hours",
        time_dbl >= min(bus_hours) | time_dbl < max(bus_hours) ~ "Business Hours"
      ) %>%
        fct(levels = c("Horrible", "Outside Hours", "Business Hours"))
    ) %>%
    ggplot(aes(month, member)) +
    geom_raster(aes(fill = status), alpha = 0.6) +
    geom_text(aes(label = time), size = 3) +
    facet_grid(rows = vars(tz), scales = "free", space = "free", ) +
    labs(x = c(), y = c(), fill = "Status") +
    scale_x_discrete(expand = expansion(0, 0)) +
    scale_y_discrete(expand = expansion(0, 0)) +
    scale_fill_manual(values = c("red3", "yellow3", "green3")) +
    theme_bw() +
    theme(
      legend.position = "bottom", 
      axis.text.y = element_markdown(), 
      strip.text.y = element_text(angle = 0),
      panel.spacing = unit(0, "npc"),
      panel.grid = element_blank()
    )
}
```


## Introduction

The initial proposal to shift meeting times in a rotating fashion was an attempt to be fair to all members, however, in reality there is a growing sense that the idea is not working as successfully as hoped for.
In reality, there will be no solution that works for everyone as the spread of time zones is just to diverse.
This is obviously an excellent problem to be wrangling, but it is still a significant challenge.
A set of proposed options are given below.

I've considered my understanding of standard business hours (8am to 5pm) to be the optimal time range for people, and between the hours of 11pm to 5am to be 'horrible'.
Anything between these is 'Outside of Hours', which should still be realistic times for everyone, despite them not being super-convenient.

An important consideration for each option is not only our own preferences, but the availability of Lori as both secretary and member of the Bioconductor Core Team, and the two co-chairs Janani & Stevie.

There are four proposals given below as the probable best options for the current CAB.

1. Placing the 'horrible times' in the empty space between the furthest spread members
2. Fixing the start time at 9am US/Eastern for every meeting
3. Starting based on the Western-most member of each region (mostly)
4. The existing schedule I can see in my calendar

I also note that time-zones may not describe exactly where everyone lives, but these are chosen to match the output of `OlsonNames()` which is how `R` wrangles time-zones.
All projections have been made based on UTC time.

Feel free to play with the code in the accompanying rmarkdown. 
I kinda had fun...


\clearpage

## CAB Description

```{r people-dates}
people <- list(
  Stevie = "Australia/Adelaide",
  Janani = "US/Mountain",
  Lori = "US/Eastern",
  Enis = "US/Eastern",
  Umar = "Africa/Lagos",
  Zahraa = "US/Central",
  Aedin = "Europe/Dublin",
  Jasmine = "US/Eastern",
  Xueyi = "Australia/Melbourne",
  Maria = "Europe/Dublin",
  Leo = "Europe/Helsinki",
  Mengbo = "Australia/Melbourne",
  Kozo = "Japan",
  Tobi = "US/Eastern",
  Nicole = "US/Pacific",
  Johannes = "Europe/Rome",
  Lluís = "Europe/Madrid",
  Kevin = "Europe/London",
  Luyi = "Asia/Shanghai",
  Hedia = "Europe/Madrid",
  Jiefei = "US/Central"
)
dates <- tibble(
  date = ymd("20250101") + seq(0, 365),
  weekday = wday(date, label = TRUE),
  month = month(date, label = TRUE),
  exists = TRUE
) %>%
  mutate(
    nbr = cumsum(exists), .by = c(month, weekday)
  ) %>%
  dplyr::filter(
    weekday == "Thu", nbr == 2, month > "Jan"
  ) %>%
  droplevels() %>% 
  dplyr::select(date, month)
```


As it currently stands, the CAB contains `r length(people)` members, spread across `r length(unique(unlist(people)))` time-zones, with these broken down and grouped in the following table.

```{r cab-summary-tbl}
people %>% 
  unlist() %>% 
  enframe(name = "member", value = "tz") %>% 
  mutate(
    current_utc = tz %>% 
      lapply(
        \(x) now(tzone = x)
      ) %>% 
      map_chr(format, "%z"),
    future_utc = tz %>% 
      lapply(
        \(x) time_at_tz(now() + 86400 * floor(365/2), x)
      ) %>% 
      map_chr(format, "%z"), 
  ) %>% 
  arrange(as.numeric(current_utc)) %>% 
  chop(member) %>% 
  mutate(
    n = map_int(member, length),
    member = map_chr(member, paste, collapse = ", ")
  ) %>% 
  rename_with(
    \(x) x %>% str_to_title() %>% str_replace_all("_utc", " UTC") %>% str_replace("Tz", "TZ")
  ) %>% 
  pander(
    justify = "lrrlr",
    caption = "
    Breakdown of all time zones with the current +UTC shift and where that shift will be 6 months into the future.
    The largest 'empty space' time period is between the US West Coast & Melbourne, with no members in the time-zones between the two.
    Currently this is 5hrs ($\\pm 1$ day), whilst in 6 months this will be 7 hours.
    The other large time gaps are between 1: Helsinki & Shanghai, and, 2: Dublin/London and the US East Coast.
    "
  )
```



```{r tbl-empty-space}
dates %>% 
  mutate(
    date = as_datetime(date),
    `US/Pacific` = time_at_tz(date, "US/Pacific"),
    `Australia/Melbourne` = time_at_tz(date, "Australia/Melbourne"),
    `Europe/Helsinki` = time_at_tz(date, "Europe/Helsinki"),
    `Asia/Shanghai` = time_at_tz(date, "Asia/Shanghai")
  ) %>% 
  mutate(
    `US/Pacific` = hour(`US/Pacific`) + minute(`US/Pacific`) / 60,
    `Australia/Melbourne` = hour(`Australia/Melbourne`) + minute(`Australia/Melbourne`) / 60,
    `Europe/Helsinki` = hour(`Europe/Helsinki`) + minute(`Europe/Helsinki`) / 60,
    `Asia/Shanghai` = hour(`Asia/Shanghai`) + minute(`Asia/Shanghai`) / 60,
    pacific_diff = abs(`Australia/Melbourne` - `US/Pacific`),
    contintental_diff = abs(`Asia/Shanghai` - `Europe/Helsinki`)
  ) %>% 
  dplyr::select(date, month, ends_with("diff")) %>% 
  mutate(
    `place_2-3am_in` = ifelse(contintental_diff == 5, "Pacific Ocean", "Helsinki-Asia")
  ) %>% 
  rename_with(\(x) x %>% str_replace_all("_", " ") %>% str_to_title()) %>% 
  pander(
    justify = "llrrl",
    caption = "
    Taking the two largest 'empty-space' regions, where no participants live, 
    and looking for the time differences.
    Time-zone differences between the US/Pacific & Melbourne are given as 'Pacific Diff', 
    whilst those between Europe/Helsinki and Asia/Shanghai are given as 'Contintental Diff'.
    "
  )
```
\clearpage

## Proposal One: Finding Empty Space

The strategy attempted here is to find the 'empty' region on the planet at each given month, where no CAB members are located, which contains the longest duration relative to *everyone's lived time-zone*.
If we place the difficult middle-of-the-night meeting times in these empty regions, the hoped for outcome will be the fewest CAB members having to endure middle-of-the-night meetings.
Of course, the CAB contains a mix of 'morning people' and 'night-owls', making this a a near impossible challenge, however, if we assume somewhere between 2-3am should be placed in the centre of these empty regions, this might be the best starting point.

For April to October, the best region to place the middle-of-the-night times is in the Pacific Ocean.
For the other times, the best region to place these is between the most eastern European time-zone (Helsinki) and western Asian time-zone (Shanghai).


```{r make-prop1-schedule}
p <- dates %>% 
  mutate(
    utc = case_when(
      month %in% c("Feb", "Mar", "Nov", "Dec") ~ as_datetime(date + 22/24),
      TRUE ~ as_datetime(date + 13/24)
    )
  ) %>% 
  draw_schedule(people)
```


Key points:

- No-one has a meeting starting anywhere between 1am and 5am
- Midnight meetings are minimised:
    - Unfortunately Leo would still have `r p$data %>% dplyr::filter(member == "Leo" & time == "00:00") %>% nrow()`
    - Mengbo & Xueyi have `r p$data %>% dplyr::filter(member == "Mengbo" & time == "00:00") %>% nrow()`
- People near the edges of the empty-space are disadvantaged
    - Most of Europe/Africa and China/Japan are inconvenient for Feb, March, Nov, Dec
    - US/Pacific, Australia/Melbourne and Japan are inconvenient for Apr to Oct
- Convenient for Lori, Janani and Stevie all year


```{r tbl-horrible}
p$data %>% 
  summarise(
    `After 11pm` = sum(time == "00:00" | time_dbl >= 23),
    `6am Or Earlier` = sum(time_dbl <= 6 & time_dbl > 0),
    .by = member
  ) %>% 
  dplyr::filter(
    rowSums(.[-1]) > 0
  ) %>% 
  bind_rows(
    summarise(.[-1], across(everything(), sum)) %>% mutate(member = "Total")
  ) %>% 
  mutate(
    member = str_remove_all(member, "\\*"),
    TZ = unlist(people)[member],
    `Combined Problematic` = `After 11pm` + `6am Or Earlier` 
  ) %>% 
  dplyr::select(Member = member, TZ, everything()) %>% 
  pander(
    justify = "llrrr",
    emphasize.strong.rows = which(str_detect(.$Member, "Total")),
    caption = "
    Summary of members with the most problematic meeting times,
    even after avoiding the 1am to 5am timezone. 
    These members are the most disadvantaged by this proposal
    "
  )
```

```{r schedule-empty-spaces, fig.cap = "Proposed schedule placing the largest time regions without members into the middle of the night"}
p
```

```{r export-empty, echo=FALSE}
write_rds(p, here::here("data", "empty_space.rds"), compress = "gz")
```


\clearpage

## Proposal Two: Fix Times at 9am For US/Eastern

Key Points:

- This is good for the key members Lori, Janani & Stevie
- **Strongly disadvantages Mengbo & Xueyi**
- Possibly workable but also inconvenient for Luyi, Kozo & Nicole

```{r schedule-ny-fixed, fig.cap = "Proposed meeting times holding the start time fixed at 9am for US/Eastern"}
p <- dates %>% 
  mutate(
    utc = case_when(
      month %in% c("Feb", "Nov", "Dec") ~ as_datetime(date + 14/24),
      TRUE ~ as_datetime(date + 13/24)
    )
  ) %>% 
  draw_schedule(people)
p
```

\clearpage

```{r tbl-horrible2}
p$data %>% 
  summarise(
    `After 11pm` = sum(time == "00:00" | time_dbl >= 23),
    `6am Or Earlier` = sum(time_dbl <= 6 & time_dbl > 0),
    .by = member
  ) %>% 
  dplyr::filter(
    rowSums(.[-1]) > 0
  ) %>% 
  bind_rows(
    summarise(.[-1], across(everything(), sum)) %>% mutate(member = "Total")
  ) %>% 
  mutate(
    member = str_remove_all(member, "\\*"),
    TZ = unlist(people)[member],
    `Combined Problematic` = `After 11pm` + `6am Or Earlier` 
  ) %>% 
  dplyr::select(Member = member, TZ, everything()) %>% 
  pander(
    justify = "llrrr",
    emphasize.strong.rows = which(str_detect(.$Member, "Total")),
    caption = "
    Summary of members with the most problematic meeting times,
    if holding the start time fixed for US/Eastern. 
    These members are the most disadvantaged by this proposal
    "
  )
```

```{r export-ny-fixed, echo=FALSE}
write_rds(p, here::here("data", "ny_fixed.rds"), compress = "gz")
```

\clearpage

## Proposal Three: Start Times Based On The Western-Most Member

This is a *slightly modified version of the current schedule*, taking the western-most, regularly attending member from each of Asia, Europe & the US as the key member.
We then start meeting times *at 7am for that member* ensuring at least two larger regions have workable times.

- 7am starts for Janani (US). These can be negotiated if Nicole is intending to resume participation in 2025 or if Janani prefers
- 7am starts for Kozo (Asia). These could also be renegotiated if Luyi is intending to resume participation in 2024, or if Kozo prefers
- From experience, most US attendees will struggle to attend after 10pm. For meetings which should be based on the Western-most European member, **these have instead been started at 6am US/Eastern** to try and bring this US/Eastern back into possibility
- Everyone gets at least 4 meetings in standard business hours (8am to 5pm as shown in green), except Luyi who has not attended a meeting on over 12 months, and Kozo
- No-one has more than 4 'horrible' meeting times (11pm to 5am)
- Janani & Kozo appear the most disadvantaged 
- For May to August, it's conceivable for the entire board to attend. Likewise, this could be repeated for Sep if this configuration was extended

```{r schedule-western, fig.cap = "Modified version of the current meeting schedule, setting the start time at 7am for the western-most member of each region (except Europe). 1) Feb-Apr favour Europe & Asia, 2) May-Aug favour the US & Europe, whilst 3) Sep-Dec favours Asia & the US"}
p <- dates %>% 
  mutate(
    utc = case_when(
      month %in% c("Feb") ~ as_datetime(date + 11/24),
      month %in% c("Mar", "Apr") ~ as_datetime(date + 10/24),
      month %in% c("May", "Jun", "Jul", "Aug") ~ as_datetime(date + 13/24),
      # month %in% c("Sep") ~ as_datetime(date + 13/24),
      month %in% c("Sep", "Oct") ~ as_datetime(date + 22/24),
      TRUE ~ as_datetime(date + 22/24)
    )
  ) %>% 
  draw_schedule(people)
p
```


```{r tbl-horrible3}
p$data %>% 
  summarise(
    `After 11pm` = sum(time == "00:00" | time_dbl >= 23),
    `6am Or Earlier` = sum(time_dbl <= 6 & time_dbl > 0),
    .by = member
  ) %>% 
  dplyr::filter(
    rowSums(.[-1]) > 0
  ) %>% 
  bind_rows(
    summarise(.[-1], across(everything(), sum)) %>% mutate(member = "Total")
  ) %>% 
  mutate(
    member = str_remove_all(member, "\\*"),
    TZ = unlist(people)[member],
    `Combined Problematic` = `After 11pm` + `6am Or Earlier` 
  ) %>% 
  dplyr::select(Member = member, TZ, everything()) %>% 
  pander(
    justify = "llrrr",
    emphasize.strong.rows = which(str_detect(.$Member, "Total")),
    caption = "
    Summary of members with the most problematic meeting times,
    if holding the start time fixed for the most western member of each region. 
    These members are the most disadvantaged by this proposal
    "
  )
```

```{r export-western, echo=FALSE}
write_rds(p, here::here("data", "western.rds"), compress = "gz")
```

\clearpage

## Proposal Four: Current Schedule

Key points:

- March & April meetings will be very difficult for Lori & Janani
- March & April will also cause problems for many in Europe & the US


```{r schedule-current, fig.cap = "The current meeting schedule, shifting by 8hrs every 4 months. Shifts have been estimated based on my own calendar invitations, so hopefully they're accurate"}
p <- dates %>% 
  mutate(
    utc = case_when(
      month %in% c("Mar") ~ as_datetime(date + 5/24),
      month %in% c("Feb", "Apr") ~ as_datetime(date + 6/24),
      month %in% c("May", "Jun", "Jul", "Aug") ~ as_datetime(date + 14/24),
      month %in% c("Nov") ~ as_datetime(date + 23/24),
      TRUE ~ as_datetime(date + 22/24)
    )
  ) %>% 
  draw_schedule(people)
p
```

\clearpage

```{r tbl-horrible4}
p$data %>% 
  summarise(
    `After 11pm` = sum(time == "00:00" | time_dbl >= 23),
    `6am Or Earlier` = sum(time_dbl <= 6 & time_dbl > 0),
    .by = member
  ) %>% 
  dplyr::filter(
    rowSums(.[-1]) > 0
  ) %>% 
  bind_rows(
    summarise(.[-1], across(everything(), sum)) %>% mutate(member = "Total")
  ) %>% 
  mutate(
    member = str_remove_all(member, "\\*"),
    TZ = unlist(people)[member],
    `Combined Problematic` = `After 11pm` + `6am Or Earlier` 
  ) %>% 
  dplyr::select(Member = member, TZ, everything()) %>% 
  pander(
    justify = "llrrr",
    emphasize.strong.rows = which(str_detect(.$Member, "Total")),
    caption = "
    Summary of members with the most problematic meeting times if adhering to the current schedule. 
    "
  )
```

```{r export-current, echo=FALSE}
write_rds(p, here::here("data", "current.rds"), compress = "gz")
```

