---
title: "Proposed CAB Meeting Times"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE,
  fig.height = 9, fig.width = 10
)
```

```{r packages}
library(tidyverse)
library(ggtext)
library(reactable)
library(htmltools)
theme_set(
  theme_bw() + theme(title = element_text(hjust = 0.5))
)
```

```{r people}
people <- list(
  Stevie = "Australia/Adelaide",
  Janani = "US/Mountain",
  Lori = "US/Eastern",
  Enis = "US/Eastern",
  Umar = "Africa/Lagos",
  Zahraa = "US/Central",
  Aedin = "Europe/Dublin",
  Jasmine = "US/Eastern",
  Xueyi = "Australia/Melbourne",
  Maria = "Europe/Dublin",
  Leo = "Europe/Helsinki",
  Mengbo = "Australia/Melbourne",
  Kozo = "Japan",
  Tobi = "US/Eastern",
  Nicole = "US/Pacific",
  Johannes = "Europe/Rome",
  Lluís = "Europe/Madrid",
  Kevin = "Europe/London",
  Luyi = "Asia/Shanghai",
  Hedia = "Europe/Madrid",
  Jiefei = "US/Central"
)
```


## Introduction

This contains a set of proposals for possible CAB meetings times to be voted on.
The full explanations are [here](../analysis/cab_meeting_time_proposal.pdf).

Meeting times for members are classified as:

1. Business Hours: 8am to 5pm meeting times
2. Outside Hours: 6-7am or 6-10pm meeting times
3. Horrible: 11pm to 5am meeting times

These can be considered overall as a reasonable description, despite some members finding a midnight meeting as quite convenient and a 6am start to be horrifying, with others being completely the opposite.

## Current Schedule {.tabset}


```{r load-current}
current <- here::here("data", "current.rds") %>% read_rds()
```

This schedule was devised by shifting the meeting in 8 hour increments every 4 months.

### Visualisation

```{r fig-current}
current + ggtitle("The Current Schedule") + theme(plot.title = element_text(hjust = 0.5))
```

### Summary Table

```{r tbl-current, echo = FALSE}
cp <- tags$caption(
  tags$em(
    "Summary of meeting times for all members under the current schedule. Members are disadvantaged in an approximately equal fashion."
  )
)
tbl <- current$data %>%
  summarise(
    n = dplyr::n(), .by = c(member, status)
  ) %>% 
  pivot_wider(names_from = "status", values_from = "n", values_fill = 0) %>% 
  mutate(
    member = fct_relabel(member, \(x) str_remove_all(x, "\\*")),
    TZ = unlist(people)[as.character(member)]
  ) %>% 
  chop(member) %>% 
  mutate(n = map_int(member, length)) %>% 
  dplyr::select(TZ, n, member, `Business Hours`, everything()) %>% 
  reactable(
    showPageSizeOptions = FALSE, filterable = TRUE,pagination = FALSE,
    columns = list(
      n = colDef(
        "In Time-Zone", maxWidth = 100, 
        footer = JS(
          "function(colInfo) {
            let n_members = 0
            colInfo.data.forEach(function(row) {
              n_members += row['n']
            })
            return (n_members).toFixed(0).bold()
          }"
        ),
        html = TRUE
      ),
      member = colDef("CAB Member(s)"),
      TZ = colDef("Time-Zone",  footer = tags$b("Total")),
      `Business Hours` = colDef(
        footer = JS(
          "function(colInfo) {
            let n_members = 0
            let hrs = 0
            let total = 0
            colInfo.data.forEach(function(row) {
              n_members = row['n']
              hrs = row['Business Hours']
              total += n_members * hrs
            })
            return (total).toFixed(0).bold()
          }"
        ),
        html = TRUE
      ),
      `Outside Hours` = colDef(
        footer = JS(
          "function(colInfo) {
            let n_members = 0
            let hrs = 0
            let total = 0
            colInfo.data.forEach(function(row) {
              n_members = row['n']
              hrs = row['Outside Hours']
              total += n_members * hrs
            })
            return (total).toFixed(0).bold()
          }"
        ),
        html = TRUE
      ),
      `Horrible` = colDef(
        footer = JS(
          "function(colInfo) {
            let n_members = 0
            let hrs = 0
            let total = 0
            colInfo.data.forEach(function(row) {
              n_members = row['n']
              hrs = row['Horrible']
              total += n_members * hrs
            })
            return (total).toFixed(0).bold()
          }"
        ),
        html = TRUE        
      )
      ),
      theme = reactableTheme(style = list(fontSize = 11)),
      elementId = "tbl-current"
  )
browsable(tagList(cp, tbl))
```

## Fixed At 9am US/Eastern {.tabset}


```{r load-ny-fixed}
ny_fixed <- here::here("data", "ny_fixed.rds") %>% read_rds()
```

This proposal attempts to help Lori continually attend, as member of the Bioconductor Core Team and CAB secretary, by starting at 9am US/Eastern every month.
For some of the year, this potentially enables full attendance.

### Visualisation

```{r fig-ny-fixed}
ny_fixed + ggtitle("Start Times Constantly at 9am (US/Eastern)") + theme(plot.title = element_text(hjust = 0.5))
```

### Summary Table

```{r tbl-ny-fixed, echo = FALSE}
cp <- tags$caption(
  tags$em(
    "Summary of meeting times for all members under the schedule fixing start times at 9am US/Eastern. Members in the Asia/Pacific region are the most strongly disadvantaged."
  )
)
tbl <- ny_fixed$data %>%
  summarise(
    n = dplyr::n(), .by = c(member, status)
  ) %>% 
  pivot_wider(names_from = "status", values_from = "n", values_fill = 0) %>% 
  mutate(
    member = fct_relabel(member, \(x) str_remove_all(x, "\\*")),
    TZ = unlist(people)[as.character(member)]
  ) %>% 
  chop(member) %>% 
  mutate(n = map_int(member, length)) %>% 
  dplyr::select(TZ, n, member, `Business Hours`, everything()) %>% 
  reactable(
    showPageSizeOptions = FALSE, filterable = TRUE,pagination = FALSE,
    columns = list(
      n = colDef(
        "In Time-Zone", maxWidth = 100, 
        footer = JS(
          "function(colInfo) {
            let n_members = 0
            colInfo.data.forEach(function(row) {
              n_members += row['n']
            })
            return (n_members).toFixed(0).bold()
          }"
        ),
        html = TRUE
      ),
      member = colDef("CAB Member(s)"),
      TZ = colDef("Time-Zone",  footer = tags$b("Total")),
      `Business Hours` = colDef(
        footer = JS(
          "function(colInfo) {
            let n_members = 0
            let hrs = 0
            let total = 0
            colInfo.data.forEach(function(row) {
              n_members = row['n']
              hrs = row['Business Hours']
              total += n_members * hrs
            })
            return (total).toFixed(0).bold()
          }"
        ),
        html = TRUE
      ),
      `Outside Hours` = colDef(
        footer = JS(
          "function(colInfo) {
            let n_members = 0
            let hrs = 0
            let total = 0
            colInfo.data.forEach(function(row) {
              n_members = row['n']
              hrs = row['Outside Hours']
              total += n_members * hrs
            })
            return (total).toFixed(0).bold()
          }"
        ),
        html = TRUE
      ),
      `Horrible` = colDef(
        footer = JS(
          "function(colInfo) {
            let n_members = 0
            let hrs = 0
            let total = 0
            colInfo.data.forEach(function(row) {
              n_members = row['n']
              hrs = row['Horrible']
              total += n_members * hrs
            })
            return (total).toFixed(0).bold()
          }"
        ),
        html = TRUE        
      )
      ),
      theme = reactableTheme(style = list(fontSize = 11)),
      elementId = "tbl-ny_fixed"
  )
browsable(tagList(cp, tbl))
```

## Setting Start Times By The Western-Most Members {.tabset}


```{r load-western}
western <- here::here("data", "western.rds") %>% read_rds()
```

This proposal shifts times every 4 months placing the western-most member of each region as the earliest start time (6am).
Until March, US/Eastern was chosen instead of Europe/Dublin as an attempt to slightly maximise attendance, given most US members find meetings after 10pm to be difficult to attend.

### Visualisation

```{r fig-western}
western + ggtitle("Starting at 6am By Region") + theme(plot.title = element_text(hjust = 0.5))
```

### Summary Table

```{r tbl-western, echo = FALSE}
cp <- tags$caption(
  tags$em(
    "Summary of meeting times starting at 6am for the most western member of each region. 
    Members on the edges of each region are the most strongly disadvantaged."
  )
)
tbl <- western$data %>%
  summarise(
    n = dplyr::n(), .by = c(member, status)
  ) %>% 
  pivot_wider(names_from = "status", values_from = "n", values_fill = 0) %>% 
  mutate(
    member = fct_relabel(member, \(x) str_remove_all(x, "\\*")),
    TZ = unlist(people)[as.character(member)]
  ) %>% 
  chop(member) %>% 
  mutate(n = map_int(member, length)) %>% 
  dplyr::select(TZ, n, member, `Business Hours`, everything()) %>% 
  reactable(
    showPageSizeOptions = FALSE, filterable = TRUE,pagination = FALSE,
    columns = list(
      n = colDef(
        "In Time-Zone", maxWidth = 100, 
        footer = JS(
          "function(colInfo) {
            let n_members = 0
            colInfo.data.forEach(function(row) {
              n_members += row['n']
            })
            return (n_members).toFixed(0).bold()
          }"
        ),
        html = TRUE
      ),
      member = colDef("CAB Member(s)"),
      TZ = colDef("Time-Zone",  footer = tags$b("Total")),
      `Business Hours` = colDef(
        footer = JS(
          "function(colInfo) {
            let n_members = 0
            let hrs = 0
            let total = 0
            colInfo.data.forEach(function(row) {
              n_members = row['n']
              hrs = row['Business Hours']
              total += n_members * hrs
            })
            return (total).toFixed(0).bold()
          }"
        ),
        html = TRUE
      ),
      `Outside Hours` = colDef(
        footer = JS(
          "function(colInfo) {
            let n_members = 0
            let hrs = 0
            let total = 0
            colInfo.data.forEach(function(row) {
              n_members = row['n']
              hrs = row['Outside Hours']
              total += n_members * hrs
            })
            return (total).toFixed(0).bold()
          }"
        ),
        html = TRUE
      ),
      `Horrible` = colDef(
        footer = JS(
          "function(colInfo) {
            let n_members = 0
            let hrs = 0
            let total = 0
            colInfo.data.forEach(function(row) {
              n_members = row['n']
              hrs = row['Horrible']
              total += n_members * hrs
            })
            return (total).toFixed(0).bold()
          }"
        ),
        html = TRUE        
      )
      ),
      theme = reactableTheme(style = list(fontSize = 11)),
      elementId = "tbl-western"
  )
browsable(tagList(cp, tbl))
```


## Difficult Meeting Times in Empty Time-Zones {.tabset}


```{r load-empty-space}
empty_space <- here::here("data", "empty_space.rds") %>% read_rds()
```

This proposal attempts to place meetings which might start between 1 and 5am either in the Pacific, between Australia & the US, or between Eastern Europe and Western Asia.

### Visualisation

```{r fig-empty-space}
empty_space + ggtitle("Difficult Meeting Start Times In Empty Time Zones") + theme(plot.title = element_text(hjust = 0.5))
```

### Summary Table

```{r tbl-empty-space, echo = FALSE}
cp <- tags$caption(
  tags$em(
    "Summary of meeting times for all members under the schedule placing the most difficult meeting times in time-zones without CAB members. 
    Members on the edges of these regions are the most strongly disadvantaged."
  )
)
tbl <- empty_space$data %>%
  summarise(
    n = dplyr::n(), .by = c(member, status)
  ) %>% 
  pivot_wider(names_from = "status", values_from = "n", values_fill = 0) %>% 
  mutate(
    member = fct_relabel(member, \(x) str_remove_all(x, "\\*")),
    TZ = unlist(people)[as.character(member)]
  ) %>% 
  chop(member) %>% 
  mutate(n = map_int(member, length)) %>% 
  dplyr::select(TZ, n, member, `Business Hours`, everything()) %>% 
  reactable(
    showPageSizeOptions = FALSE, filterable = TRUE,pagination = FALSE,
    columns = list(
      n = colDef(
        "In Time-Zone", maxWidth = 100, 
        footer = JS(
          "function(colInfo) {
            let n_members = 0
            colInfo.data.forEach(function(row) {
              n_members += row['n']
            })
            return (n_members).toFixed(0).bold()
          }"
        ),
        html = TRUE
      ),
      member = colDef("CAB Member(s)"),
      TZ = colDef("Time-Zone",  footer = tags$b("Total")),
      `Business Hours` = colDef(
        footer = JS(
          "function(colInfo) {
            let n_members = 0
            let hrs = 0
            let total = 0
            colInfo.data.forEach(function(row) {
              n_members = row['n']
              hrs = row['Business Hours']
              total += n_members * hrs
            })
            return (total).toFixed(0).bold()
          }"
        ),
        html = TRUE
      ),
      `Outside Hours` = colDef(
        footer = JS(
          "function(colInfo) {
            let n_members = 0
            let hrs = 0
            let total = 0
            colInfo.data.forEach(function(row) {
              n_members = row['n']
              hrs = row['Outside Hours']
              total += n_members * hrs
            })
            return (total).toFixed(0).bold()
          }"
        ),
        html = TRUE
      ),
      `Horrible` = colDef(
        footer = JS(
          "function(colInfo) {
            let n_members = 0
            let hrs = 0
            let total = 0
            colInfo.data.forEach(function(row) {
              n_members = row['n']
              hrs = row['Horrible']
              total += n_members * hrs
            })
            return (total).toFixed(0).bold()
          }"
        ),
        html = TRUE        
      )
      ),
      theme = reactableTheme(style = list(fontSize = 11)),
      elementId = "tbl-empty_space"
  )
browsable(tagList(cp, tbl))
```

